An optimization of bidirectional update language based on the idea of reversible computing \thanks{Supported by organization x.}

### 2019-09-08 paper/main.tex
only update formulas !

\textcolor{blue}{(should substitute a, b, c, d, e, f by s1, s2, v1, v2...)}\\
$\pg{bx_1 \circ bx_2}{s}{v} \\
    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tab = a \Leftarrow \getbx{bx_1}{s}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
            \qtab \tab (d, c) \\
    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
            \qtab \tab (d, c) \\
    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
        \qtab (d, e) \Leftarrow \pg{bx_1}{f}{b}; \\
            \qtab \tab (d, c)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $a \Leftarrow \getbx{bx_1}{s}$ into $(f, a) \Leftarrow \pg{bx_1}{s}{dummy}$. It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid. The last equality substitutes $s$ in $(d, e) \Leftarrow \pg{bx_1}{s}{b}$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $e$ in the result pair $(d,e)$ equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
    \tab \text{if} \ cond_{sv} \ s \ v\\
    \tab \text{then} \ (ks, kv, s, v) \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
    \tab \text{else} \ (ks, kv, s, v) \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
    \tab \text{fi} \ cond_{s} \ s; \ \text{return} \ (ks, kv, s, v)$\\

$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{const$\_$dummy \ v};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$\\

@@

\textcolor{blue}{(how to explain parameters of $cpg$ in the composition?)}

@@

    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
        \qtab (s; v)$

\textcolor{blue}{(some explanation for parameters $\lambda \_.s, \lambda \_.v$)}

@@

\textcolor{blue}{(some explanation for parameters $\lambda \_.s, id, id, id$)}\\

@@

$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
    \tab \text{if} \ cond_{sv} \ s \ v \\
    \tab \text{then} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{else} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{fi} \ cond_s \ (ks' \ s); \ \text{return} \ (ks, kv, ks', kv', s, v)$\\

$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{const$\_$dummy \ v};\\
    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
    \qtab (ks_1 \circ ks_2, kv_2', ks_1 \circ ks_2', kv_2', s_2, v_2)$\\

@@

$\xpg{bx_1 \circ bx_2}{s}{v} =\\
    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{v};\\
    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
    \qtab (ks_1 \ s_2, v_2)$

@@

In section 5, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view. Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s. For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ direction on the composition valid.\\

>>>>

$\pg{bx_1 \circ bx_2}{s}{v} \\
    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
            \qtab \tab (s_3, v_2) \\
    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
            \qtab \tab (s_3, v_2) \\
    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tab (s_3, v_2)$\\

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid. The last equality substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$. This transformation is feasible due to the PUTPUT law. Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$. So we can realize that there is no loss information when computing a composition.

@@

$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
    \tab \text{fi} \ cond_{s} \ s' \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\

$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$\\

@@

@@

    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
        \qtab (s; v)$

@@

@@

$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{fi} \ cond_s \ (ks' \ s') \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\

$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$\\

@@

$\xpg{bx_1 \circ bx_2}{s}{v} =\\
    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
    \qtab (ks_1 \ s_2, v_2)$

@@

In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view. Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s. For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ direction on the composition valid.\\
